Child survival for mothers: mortality change and related measures

### 2020-04-15 ALAP2020/ALAP.Rmd
abs change infant mortality

We now consider the consequences of a change $\delta$ in mortality in the range $[0,a-\alpha]$, where $\alpha$ is the start age of fertility risk. A discrete approximate of the effect on $CS_a$ given an absolute change in mortality, considering that $m_{x,\delta}=m_{x}+\delta$ (@wrycza2012), and $l_{a-x} = e^{-\int_{0}^{a-x}{(\mu_t+\delta})}$:

@@

@@

$$\frac{\Delta CS_{a}}{CS_a} = - \left[\int_{0}^{a}{\mu_t  \, \frac{CS_{a-t}}{CS_a} \,dt \,}\right] \Delta \delta$$
```{r main_funtion}
# age survivor spline

>>>>

We now consider the consequences of a change $\delta$ in mortality in the range $[0,a-\alpha]$, where $\alpha$ is the start age of fertility risk. A discrete approximate of the effect on $CS_a$ given an absolute change in mortality, considering that $m_{x,\delta}=m_{x}+\delta$ (@wrycza2012), and $l_{a-x} = e^{-\int_{0}^{a-x}{(\mu_t+\delta})dt}$:

@@

If the change is only in the age range $\left[0;1\right)$, infant mortality, we can start by splitting the integral^[$a>1$]:

$$CS_{a}^{\delta_0} = \int_{0}^{a-1}
m_{x} {e{^{-\int_{1}^{a-x}\mu_t \,dt-\int_{0}^{1}(\mu_t + \delta)\,dt}}} dx + 
\int_{a-1}^{a} m_{x} e^{-\int_{0}^{a-x}(\mu_t+ \delta)\,dt}\\
=\int_{0}^{a-1} m_{x} l_{a-x}e^{-\delta} dx + \int_{a-1}^{a} m_{x} l_{a-x} e^{-\delta(a-x)}\,dx
$$

Deriving by $\delta$ and valuating near 0, we get:

$$CS_{a}^{\delta_0} =-\int_{0}^{a-1} m_{x} l_{a-x} dx - a \int_{a-1}^{a} {m_{x} l_{a-x} dx} + \int_{a-1}^{a} {x \, m_{x} l_{a-x} dx}$$
Expressing $\int_{0}^{a-1} m_{x} l_{a-x} dx$ as $CS_a - CS_{a-1,a}$, and finally dividing by $CS_a$:

$$ = - CS_a + CS_{a-1,a} (1-a+\kappa_{a-1,a})\\
\frac{\Delta CS_{a}}{CS_{a}} \approx - \left[1-\frac{CS_{a-1,a}}{CS_a} (1-(a-\kappa_{a-1,a}))\right] \Delta\delta$$

Affects proportionally for the range extension of 1, less $P_b$ (if we assume uniform distribution of deaths and births in the first year) times $\frac{CS_{a-1,a}}{CS_a}$ the portion of cumulated alive daughters in the previous year, weighting how important was the last year of reproduction.

@@

$$\frac{\Delta CS_{a}}{CS_a} = - \left[\int_{0}^{a}{\mu_t  \, \frac{CS_{a-t}}{CS_a} \,dt \,}\right] \Delta \delta$$

In a model with heterogeneity, determined at born with a multiplicative effect, $CS_a$ could be interpreted as a conditional expectation with random variables *Y* for fertility heterogeneity and *Z* for frailty.

$$CS_a = \int_{0}^{\infty}f_y \int_{0}^{\infty}f_{z|y}\left[\int_{0}^{\infty}y\,\overline{m}_x\,e^{-\overline{H}_{a-x}\,z}\,dx\right]\,dz\,dy\\
CS_a = \int_{0}^{\infty}f_y \, y \int_{0}^{\infty}f_{z|y}\left[\int_{0}^{\infty}\,\overline{m}_x\,e^{-\overline{H}_{a-x}\,z}\,dx\right]\,dz\,dy
$$

```{r main_funtion}
# age survivor spline
